\documentclass[11pt]{article}
\setlength{\parindent}{0. true in}
%\pagestyle{empty}
\textwidth 6.25 truein
\oddsidemargin 0.1 truein
\evensidemargin .1 truein
\topmargin -.8 truein
\textheight 9.7 truein
\begin{document}

%%%%%%%%%%%% START OF GRAF BIO %%%%%%%%%%%%%%%%
%\newpage

\vskip5pt
\begin{center}
\begin{tabular}{c}
{\bf Peter Graf, NREL}\\
15013 Denver West Parkway MS ESIF301, Golden CO 80401\\
303-275-4666, peter.graf@nrel.gov\\
\hline
\end{tabular}
\end{center}


\vskip5pt


\subsubsection*{Education and Training}

\begin{tabular}{ll}

B.S., Symbolic Systems, Stanford University, 1989\\

Graduate Research Associate, Lawrence Berkeley National Laboratory, 1998-2002\\

Ph.D., Mathematics, University of California at Berkeley, 2003\\

Postdoctoral Research Associate, NREL, Scientific Computing Center, 2003-2006\\

\end{tabular}



\subsubsection*{Research and Professional Experience}

\begin{tabular}{l p{5.5in}l}

2006-- & Senior Scientist, NREL, Computational Science Center.  High performance computing.  Applied mathematics.  Modeling, simulation, optimization, data mining in support of a variety of renewable energy research.  Mathematical research in simulation optimization and multi-scale simulation.\\
2003--2006 & Postdoctoral Research Associate, NREL, Scientific Computing Center.  Simulation optimization for alloys and nanostructures.\\
1998--2002 & Graduate Research Associate, Lawrence Berkeley National Laboratory.  Mathematical research in optimal prediction and other model reduction techniques for differential equations.\\
1992--1997 & C/C++ Programmer, Access Softek, Inc. Berkeley, Ca.  Windows and Macintosh commercial programmer. Highlights from a wide variety of projects include genetic programming based optimization for physical simulation of articulated figures.\\

\end{tabular}



\subsubsection*{Selected Publications}

\begin{enumerate}
\item S. C. Billups, J. Larson, and P. Graf (2013), 
Derivative-Free Optimization of Expensive Functions with Computational Error Using Weighted Regression,
SIAM J. Optim., Vol. 23 (1), 27Ð53.

\item A. M. Nardes, A. J. Ferguson, J. B. Whitaker, B. W. Larson, R. E.
Larsen, K. Maturov\'a, P. A. Graf, O. V. Boltalina, S. H. Strauss, N.
Kopidakis, (2012), Beyond PCBM: Understanding the photovoltaic performance of blends of indene-C60
multiadducts with poly(3-hexylthiophene), 
Advanced Functional Materials
Vol. 22, (19), pp. 4115Ð4127.

\item M. Lunacek, A. Nag, D. A. Alber, K. Gruchalla, C. Chang, and P. A. Graf, (2011), Simulation, Characterization, and Optimization of Metabolic Models with the High Performance Systems Biology Toolkit, SIAM J. Sci. Comput. Vol. 33, pp. 3402-3424.

%\item D. Biagioni, D. P. Astling, P. A. Graf, and M. F. Davis (2011), Orthogonal projection to latent structures solution properties for chemometrics and systems biology data, J. Chemometrics, Vol. 25, 9.

\item M. E. K\"ose, P. A. Graf, N. Kopidakis, S. Shaheen, K. Kim, and G. Rumbles, (2009), Exciton Migration in Conjugated Dendrimers: A Joint Experimental and Theoretical Study, ChemPhysChem,  Vol. 10, (18), pp. 3285-3294.

\item P. Graf, M. E. K\"ose and K. Kim, Numerical Formulation of the Effective Medium Approximation: Illustrative Examples and Application to Organic Semiconductors, in Computational Nanoscience Ñ How to Exploit Synergy between Predictive Simulations and Experiment, (Mater. Res. Soc. Symp. Proc. Volume 1177E, Warrendale, PA, 2009), Z7.

\item P. A. Graf, W. B. Jones, and K. Kim, (2009), A note on the virtual crystal approach to alloy optimization, J. Comput. Phys., Vol. 228, pp. 4309.

\item   P. Piquini, P. A. Graf, and A. Zunger, (2008), Band-gap Design of Quaternary (In,Ga)(As,Sb) Semiconductors Via the Inverse-band-structure Approach, Phys. Rev. Lett. Vol. 100, 2008; pp. 186403.

\item P. A. Graf and W. B. Jones (2007), A Projection Based Multiscale Optimization Method for Eigenvalue Problems. J. Global Optim. Vol 39, 2007; pp. 235-245.

\item P. A. Graf, K. Kim, W. B. Jones, and L.-W. Wang, (2007), Surface Passivation Optimization Using DIRECT. J. Comput. Phys. Vol. 224(2), 2007; pp. 824-835.

%\item K. Kim, P. A. Graf, W. B. Jones and G. L. W. Hart, Direct enumeration of alloy configurations for electronic structural properties, (2005), Appl. Phys. Lett. Vol. 87, 243111.

\item K. Kim, P. A. Graf, and W. B. Jones, W.B. (2005), A Genetic Algorithm Based Inverse Band Structure Method for Semiconductor Alloys, J. Comput. Phys. Vol. 208(2), pp. 735-760.

\end{enumerate}



\subsubsection*{Synergistic Activities}

\begin{itemize}
\item Moderator, NREL Systems Engineering for Wind Energy Workshop, 2013

\item  Reviewer, ASCR Early Career Research Program, 2011

\item Reviewer, SuperComputing08 Technical Program Committee

\item Reviewer, SuperComputing09 Technical Program Committee

\end{itemize}



\subsubsection*{Collaborators and Co-Editors (past 48 months)}

David Biagioni (CU-Boulder),
Stephen Billups (CU-Denver), 
Katherine Dykes (NREL),
Gi-Heon Kim (NREL),
Stephan Lany (NREL),
Ross Larsen (NREL), 
Monte Lunacek (CU-Boulder)
Andrew Ning (NREL),
Michael Sprague (NREL),
Sean Shaheen (Denver Univ.),
Kandler Smith (NREL),
Vladan Stevanovich (NREL),
Alex Zunger (CU-Boulder)

\subsubsection*{Graduate \& Postdoctoral Advisors}

Alexandre Chorin (UC-Berkeley), Wesley Jones (NREL), Kwiseon Kim


\subsubsection*{Graduate and Postdoctoral Advisees:}

David Alber (WalkScore), Monte Lunacek (CU-Boulder), Dave Biagioni (CU-Boulder/NREL)

%%%%%%%%%%%% END OF GRAF BIO %%%%%%%%%%%%%%%%
\end{document}