\documentclass[11pt]{article}
\setlength{\parindent}{0. true in}
%\pagestyle{empty}
\textwidth 6.25 truein
\oddsidemargin 0.1 truein
\evensidemargin .1 truein
\topmargin -.8 truein
\textheight 9.7 truein
\begin{document}


\begin{center}
{\large{\bf George Pau -- Conflict of Interest}} \\
\end{center}

\vspace{.2 true in}

Co-authors and collaborators
\begin{itemize}
\item Agarwal, Deb. Lawrence Berkeley National Laboratory.
\item Almgren, Ann S. Lawrence Berkeley National Laboratory.
\item	Cances, Eric. Ecole des Ponts Paris Tech, France.
\item	Day, Marcus. Lawrence Berkeley National Laboratory.
\item	Fagnan, Kirsten M. Lawrence Berkeley National Laboratory.
\item	Finsterle, Stefan A. Lawrence Berkeley National Laboratory.
\item	Gable, Carl W. Pacific Northwest National Laboratory./
\item	Gordon, Ian. Pacific Northwest National Laboratory.
\item	Porter, Ellen A. Pacific Northwest National Laboratory.
\item	Pruess, Karsten. Lawrence Berkeley National Laboratory.
\item	Lansing, Carina. Pacific Northwest National Laboratory.
\item	Le Bris, Claude. Ecole des Ponts ParisTech, France.
\item	Lijeswki, Michael J. Lawrence Berkeley National Laboratory.
\item	Liu, Gui-rong. University of Cincinnanti.
\item	Maday, Yvon. Universitie Paris VI - Pierre et Marie Curie, France.
\item	Nguyen, Ngoc C. Massachusetts Institute of Technology.
\item	Sivaramakrishnan, Chandrika. Pacific Northwest National Laboratory.
\item	Purohit, Sumit. Pacific Northwest National Laboratory.
\item	Gosink, Luke. Pacific Northwest National Laboratory.
\item	Schuchardt, Karen L. Pacific Northwest National Laboratory.
\item	Sonnenthal, Eric. Lawrence Berkeley National Laboratory.
\item	Spycher, Nicholas. Lawrence Berkeley National Laboratory.
\item	Vesselinov, Velimir V. Los Alamos National Laboratory.
\item	Xu, Tianfu. Lawrence Berkeley National Laboratory.
\item	Zhang, Guoxiang. Shell International Exploration \& Production Inc.
\item	Zhang, Keni. Lawrence Berkeley National Laboratory.
\item	Zhang, Yingqi. Lawrence Berkeley National Laboratory.
\item	Zheng, Hui. Institute of High Performance Computing, Singapore.
\end{itemize}


\vspace{.2 true in}
Graduate and Postdoctoral Advisors
\begin{itemize}
\item	Bell, John B. Lawrence Berkeley National Laboratory. Postdoctoral Advisor.
\item	Patera, Anthony T.  Massachusetts Institute of Technology. Graduate Advisor.
\end{itemize}

\end{document}
